Modular transmission line probes for microfluidic nuclear magnetic resonance spectroscopy and imaging

### 2019-02-07 ms5n17-tlp-mn-180216-001.tex
Ready for submission

In this paper we present a novel modular probe designed for generic
microfluidic-NMR experiments. The design is based on a transmission line
detector~\cite{stripline_jan,gream_2016} using inexpensive
printed circuit board (PCB) technology. A modular design ensures that detectors of
different geometry and size can be easily exchanged.
The NMR detector (including the tuning and matching circuit) is connected to the probe
skeleton through detachable connectors. Modularity of the probe not only allows
fast optimisation of the detector but also facilitates the use of dedicated detectors
for different microfluidic devices on a single probe base.
Once the probe base is in place, additional detectors can be manufactured
within 2-3 days for less than $\sim$200 $\geneuro$.
The probe is designed for proton detected double resonance.
The second channel can be tuned to different nuclei such as $^{13}$C, $^{15}$N, or
$^{31}$P. The probe skeleton is made almost entirely from easily obtainable raw materials
such as aluminium profiles of standard size, with only a small number of
custom-machined parts holding the structure together.
Detailed construction drawings and a complete set of CAD files is
available for download at \cite{github-probe-download-2019}.
The planar geometry of the detector accommodates exchangeable microfluidic devices
which can be designed freely for specific experimental requirements as long as the
overall shape of the device is maintained.

In the remainder of this paper, we discuss the probe design in more detail,
and characterise its performance in terms of sensitivity and resolution.
We also demonstrate its application in microfluidic cell culture,
protein NMR spectroscopy, and micro-imaging.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Probe Design}
\label{sec:probe-design}
%Figure showing modular probe concept; example of probe base with several inserts\newline
%\begin{itemize}
%	\item Philosophy: use cheap materials and prefab Al profiles wherever possible
%	\item Hold it together with a small number of precision-machined parts
%	\item Generic probe base design that can accommodate a variety of inserts
%	\item Insert: PCB that accommodates both tuning/matching circuitry for up to two channels and the transmission line probe itself
%	\item Tuning and matching capacitors for each channel are part of the probe base
%	\item Connection using standard coaxial RF connectors (SMP)
%	\item Microfluidic chip design; outline
%	\item Design/explosion drawings
%	\item Circuit design for double resonance
%	\item Transmission line connection between tuning/matching of 1H channel and resonator (Smith chart argument)
%	\item 38 mm bore (Bruker NB) and 45 mm bore (Varian/Agilent NB)
%\end{itemize}
\begin{figure}
\centering
\includegraphics[width=.5\linewidth,keepaspectratio=true]{./figures/ms5n17-tlp-im-190205-rf-device-assembly.png} 
\caption{(A) Orientation of the device with respect to the constriction on the detector (PCB plane 5 is hidden). (B) Exploded view of the rf assembly with the device. }
\label{fig:rf-device-assembly} 
\end{figure}

\begin{figure}
\centering
\includegraphics[width=.5\linewidth,keepaspectratio=true]{./figures/ms5n17-tlp-im-181218-probe.png}
\caption{The probe with and without sleeve.}
\label{fig:probe}
\end{figure}
The probe assembly design is based on readily and easily available materials and parts
with minimal need for customised fabrication.
All mechanical and electronic components are readily available from commercial sources,
with the exception of a small number of precision-machined parts which are used
to hold the assembly together. These can be ordered from commercial prototyping
manufacturers or machined in-house if mechanical workshop facilities are available.

\begin{figure}
\centering
\includegraphics[width=.6\linewidth,keepaspectratio=true]{./figures/ms5n17-tlp-im-181007-Probe-explode.png} 
\caption{Exploded view of the probe design (left) and assembled probe (right). 
Probe base (blue) and scaffold assembly (green) with TMPCB (brown, 7) provides a 
generic mounting point for the NMR detector assembly (red). 
A microfluidic device (1) can be inserted from the top of the probe. 
The device is held in place by a spacer (4) between the detector planes (3 and 5) 
ensuring alignment of the sample chamber with the sensitive area on the PCB planes.
The modular NMR detector is made up of parts 3,4 and, 5 and is 
connected to the TMPCB (7) through standard coaxial (SMP) connectors. 
The scaffold assembly is made up of aluminium tubes (11), scaffold support disks (9)
and, auxiliary support disks (12).}
\label{fig:probe-explode}
\end{figure}

The probe is composed of three main assemblies: A) probe base; B) scaffold assembly;
C) detector assembly (\fig{fig:probe}).
The  base (blue in \fig{fig:probe-explode}) and the scaffold assembly
(green in \fig{fig:probe-explode}) together make the skeleton of the probe.
Both the base and scaffold assembly are made from standard aluminium profiles.
The base contains the BNC connectors for RF (Radio Frequency) input.
These BNC connectors are connected to semi-rigid coaxial cables for
RF transport in the probe.

@@

The base, the scaffold assembly and the tuning and matching PCB all have a free circular
axial bore providing access from the bottom of the probe, which can be used for
temperature and environmental control and/or fluidic connections to the sample as
needed.

The detector assembly is made up of two PCBs (3 and 5), a spacer between them (4). The detector used here is based on a planar transmission line geometry

as described elsewhere~\cite{gream_2016,stripline_jan}; other structures such as spiral
coils and Helmholtz pairs are also possible, and can easily be realised on a PCB.
The transmission line detector consists of a pair of copper
planes, each with a constriction at the centre.  This geometry gives rise to an
electromagnetic eigenmode with a strong anti-node of the magnetic field between
the two planes at the site of the constriction. This concentrates the rf field
and the detection sensitivity
onto the sample area.
The shape and size of the copper structures have been chosen such as to produce
an eigenfrequency of this mode around 630~MHz. The PCBs have been designed with
solder points at the back side for chip capacitors; these can be used to tune the
resonance down to the desired proton Larmor frequency (500 or 600~MHz in the present case).
Both the PCBs (3 and 5) also carry circuitry for channel separation and primary
matching of the detector. A copper or brass auxiliary disk (2) holds the PCBs
together at the top, and forms the top end of the probe assembly.
The probe sheath (not shown in~\fig{fig:probe-explode}) is fixed to the auxiliary disk (2)
and the scaffold assembly with screws.
An exchangeable microfluidic device (1) can be inserted from the top of the detector.
Microfluidic devices can be used as passive sample holders, filled before putting them
into the probe,
or the device can be actively perfused and operated in-situ.
The microfluidic device is held in place by the spacer
between the two PCB planes coinciding the sample chamber with the constriction on the
detector.  The 45-degree edges of the device outline help with precise and reliable
alignment.
Since the detector assembly is attached to the probe base through SMP
connectors, it can easily be exchanged.
Using standard PCB processes to manufacture the detector provides great flexibility
to optimise detector performance for different applications, including different
microfluidic sample sizes, and different combinations of target nuclei.

@@

The double resnance circuit is based on the design shown in \fig{fig:circuit}.
%The size of the constriction was larger compared to~\cite{gream_2016} to allow use of larger
%sample volumes for some cases.
The transmission line detector is tuned to the desired proton
Larmor frequency by chip capacitors located at its four corners,
as well as a non-magnetic trimmer capacitor ($C_6$). Two primary
matching capacitors ($C_5$ and $C_5'$) bring the impedance into the
vicinity of 50$\Omega$. A short microstrip
with a specific impedance of 50$\Omega$
on the PCB then provides a connection to the secondary tuning and matching elements on the TMPCB ($C_1$ and $C_2$).
The lower frequency channel is connected to the detector through a high-frequency band reject ($C_7$, $L_7$).
Tuning is achieved by $L_1$ and $C_4$, which are both located on the exchangeable insert, in combination with
the complex admittance of the detector at the X channel frequency.
To maximise efficiency, $L_1$ should have as high a $Q$-factor as possible. We have
used commercially available flat copper coils of 5 mm diameter and 4.5 turns in the case of
\textsuperscript{15}N at 60 MHz. Tuning to \textsuperscript{13}C, either at 125 or at 150 MHz
could be achieved without $L_1$.

@@

To avoid the inhomogeneities of FR4 and therefore to achieve better resolution,
detectors were made from RO3035 material. While RO3035 provides better
resolution, it is more expensive and takes longer to obtain compared to FR4. It
therefore makes sense to first prototype new detector designs in FR4, and only
use the more expensive RO3035 once the design has been validated.

\begin{figure} \centering
\includegraphics[width=.9\linewidth,keepaspectratio=true]{./figures/pcb-fieldmap-colorbar.png}
\caption{B\textsubscript{0} field map of the sample chamber for FR4 (A) and
Rogers RO3035 (B) PCB material. Both the field maps were constructed from 2
gradient echo images acquired on a chip filled with 130 mM sodium acetate
dissolved in H\textsubscript{2}O through same parameters and pulse sequence.}
\label{fig:field map}
\end{figure}

@@

\begin{figure} \centering
\includegraphics[width=\linewidth,keepaspectratio=true]{./figures/ms5n17-tlp-sp-180612-HSQCspect.png}
\caption{ $^1$H-$^{13}$C HSQC spectrum of $^{13}$C labeled glucose (left) and
$^1$H-$^{15}$N HSQC of $^{15}$N labeled Ubiquitin (right) both acquired from
2.5$\mu$l sample volume at 14.1 T. Glucose HSQC spectrum was acquired in 12
minutes from a 100 mM sample. Ubiquitin HSQC spectrum was acquired in 400
minutes from a 1 mM sample.} \label{fig:HSQC} \end{figure}
%$
The probe can also be used to acquire micro magnetic resonance (MR) images using
a Bruker micro-gradient unit. To this effect, the aluminium probe sleeve must be
replaced by an electrically insulating material in order to avoid image
distortions due to eddy currents in the sleeve. \fig{fig:tisli} shows MR images
of mouse liver tissue slice immersed in DMEM cell growth medium. The slice was
placed in a microfluidic device with circular sample chamber centred at the
detector location, as shown in \fig{fig:tisli}(A). The image shown in
\fig{fig:tisli}(B) was acquired using a FLASH pulse sequence with $t_R=120$~ms
and $t_E=15.3$~ms. The slice thickness of  1~mm encompasses the entire thickness
of the tissue; the lateral dimensions of all MR images in \fig{fig:tisli} are
512$\times$512 pixels at a resolution of 30~$\mu$m. \fig{fig:tisli}(C) shows a
RARE image using the same parameters, with 8 echos acquired per scan, thus
reducing the acquisition time to 8 minutes. A spin echo image is shown in
\fig{fig:tisli}(D), acquired over 25~min. Finally, \fig{fig:tisli}(E) shows a
$B_0$ field map, which has been obtained by acquiring two separate FLASH images
with echo times $t_E=5.3$~ms and 15.3~ms, respectively. , corresponding to local
field variations of $\Delta B/B_0=...$ \todo{compute value and insert here}.

>>>>

In this paper we present a novel modular probe designed for generic
microfluidic-NMR experiments. The design is based on a transmission line
detector~\cite{stripline_jan,gream_2016} using inexpensive
printed circuit board (PCB) technology. A modular design ensures that detectors of
different geometry and size can be easily exchanged.
The NMR detector (including the tuning and matching circuit) is connected to the probe
skeleton through detachable connectors. Modularity of the probe not only allows
fast optimisation of the detector but also facilitates the use of dedicated detectors
for different microfluidic devices on a single probe base.
Once the probe base is in place, additional detectors can be manufactured
within 2-3 days for less than $\sim$200 $\geneuro$.
The probe is designed for proton detected double resonance.
The second channel can be tuned to different nuclei such as $^{13}$C, $^{15}$N, or
$^{31}$P. The probe skeleton is made almost entirely from easily obtainable raw materials
such as aluminium profiles of standard size, with only a small number of
custom-machined parts holding the structure together.
Detailed construction drawings and a complete set of CAD files is
available for download at \url{https://github.com/marcel-utz/modular-microfluidic-probe}.
The planar geometry of the detector accommodates exchangeable microfluidic devices
which can be designed freely for specific experimental requirements as long as the
overall shape of the device is maintained. \fig{fig:rf-device-assembly} shows the microfluidic device and the detector with their relative orientation.

In the remainder of this paper, we discuss the probe design in more detail,
and characterise its performance in terms of sensitivity and resolution.
We also demonstrate its application in microfluidic cell culture,
protein NMR spectroscopy, and micro-imaging.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Probe Design}
\label{sec:probe-design}
%Figure showing modular probe concept; example of probe base with several inserts\newline
%\begin{itemize}
%	\item Philosophy: use cheap materials and prefab Al profiles wherever possible
%	\item Hold it together with a small number of precision-machined parts
%	\item Generic probe base design that can accommodate a variety of inserts
%	\item Insert: PCB that accommodates both tuning/matching circuitry for up to two channels and the transmission line probe itself
%	\item Tuning and matching capacitors for each channel are part of the probe base
%	\item Connection using standard coaxial RF connectors (SMP)
%	\item Microfluidic chip design; outline
%	\item Design/explosion drawings
%	\item Circuit design for double resonance
%	\item Transmission line connection between tuning/matching of 1H channel and resonator (Smith chart argument)
%	\item 38 mm bore (Bruker NB) and 45 mm bore (Varian/Agilent NB)
%\end{itemize}
\begin{figure}
\centering
\includegraphics[width=.8\linewidth,keepaspectratio=true]{./figures/ms5n17-tlp-im-190205-rf-device-assembly.png} 
\caption{Drawings of the detector assembly and the microfluidic device (1) in (A) front view and, (B) exploded view. Spacer (4) ensures the alignment of the sample chamber with the constrictions on the PCB planes. In (A), PCB plane 5 is hidden to show the orientation of 1 with respect to PCB plane 3.}
\label{fig:rf-device-assembly}
\end{figure}

%\begin{figure}
%\centering
%\includegraphics[width=.5\linewidth,keepaspectratio=true]{./figures/ms5n17-tlp-im-181218-probe.png}
%\caption{The probe with and without sleeve.}
%\label{fig:probe}
%\end{figure}
The probe assembly design is based on readily and easily available materials and parts
with minimal need for customised fabrication.
All mechanical and electronic components are readily available from commercial sources,
with the exception of a small number of precision-machined parts which are used
to hold the assembly together. These can be ordered from commercial prototyping
manufacturers or machined in-house if mechanical workshop facilities are available.

\begin{figure}
\centering
\includegraphics[width=.8\linewidth,keepaspectratio=true]{./figures/ms5n17-tlp-im-181007-Probe-explode.png} 
\caption{Exploded view of the probe design (left) and assembled probe (right). 
Probe base (blue) and scaffold assembly (green) with TMPCB (brown, 7) provides a 
generic mounting point for the NMR detector assembly (red). 
A microfluidic device (1) can be inserted from the top of the probe. 
The device is held in place by a spacer (4) between the detector planes (3 and 5) 
ensuring alignment of the sample chamber with the sensitive area on the PCB planes.
The modular NMR detector is made up of parts 3,4 and, 5 and is 
connected to the TMPCB (7) through standard coaxial (SMP) connectors. 
The scaffold assembly is made up of aluminium tubes (11), scaffold support disks (9)
and, auxiliary support disks (12).}
\label{fig:probe-explode}
\end{figure}

The probe is composed of three main assemblies: A) probe base; B) scaffold assembly;
C) detector assembly (\fig{fig:probe-explode}).
The  base (blue in \fig{fig:probe-explode}) and the scaffold assembly
(green in \fig{fig:probe-explode}) together make the skeleton of the probe.
Both the base and scaffold assembly are made from standard aluminium profiles.
The base contains the BNC connectors for RF (Radio Frequency) input.
These BNC connectors are connected to semi-rigid coaxial cables for
RF transport in the probe.

@@

The base, the scaffold assembly and the tuning and matching PCB all have a free circular
axial bore providing access from the bottom of the probe, which can be used for
temperature and environmental control and/or fluidic connections to the sample as
needed. The detector assembly is made up of two PCBs (3 and 5), a spacer between them (4). The detector used here is based on a planar transmission line geometry as described elsewhere~\cite{gream_2016,stripline_jan}; other structures such as spiral
coils and Helmholtz pairs are also possible, and can easily be realised on a PCB.
The transmission line detector consists of a pair of copper
planes, each with a constriction at the centre.  This geometry gives rise to an
electromagnetic eigenmode with a strong anti-node of the magnetic field between
the two planes at the site of the constriction. This concentrates the rf field
and the detection sensitivity
onto the sample area.
The shape and size of the copper structures have been chosen such as to produce
an eigenfrequency of this mode around 630~MHz. The PCBs have been designed with
solder points at the back side for chip capacitors; these can be used to tune the
resonance down to the desired proton Larmor frequency (500 or 600~MHz in the present case).
Both the PCBs (3 and 5) also carry circuitry for channel separation and primary
matching of the detector. A copper or brass auxiliary disk (2) holds the PCBs
together at the top, and forms the top end of the probe assembly.
The probe sheath (not shown in~\fig{fig:probe-explode}) is fixed to the auxiliary disk (2)
and the scaffold assembly with screws.
An exchangeable microfluidic device (1) can be inserted from the top of the detector.
Microfluidic devices can be used as passive sample holders, filled before putting them
into the probe,
or the device can be actively perfused and operated in-situ.
The microfluidic device is held in place by the spacer
between the two PCB planes coinciding the sample chamber with the constriction on the
detector.  The 45-degree edges of the device outline help with precise and reliable
alignment.
Since the detector assembly is attached to the probe base through SMP
connectors, it can easily be exchanged.
Using standard PCB processes to manufacture the detector provides great flexibility
to optimise detector performance for different applications, including different
microfluidic sample sizes, and different combinations of target nuclei.

@@

The double resnance circuit is based on the design shown in \fig{fig:circuit}.
%The size of the constriction was larger compared to~\cite{gream_2016} to allow use of larger
%sample volumes for some cases.
The transmission line detector is tuned to the desired proton
Larmor frequency by chip capacitors located at its four corners,
as well as a non-magnetic trimmer capacitor ($C_6$). Two primary
matching capacitors ($C_5$ and $C_5'$) bring the impedance into the
vicinity of 50$\Omega$. A short microstrip
with a specific impedance of 50$\Omega$
on the PCB then provides a connection to the secondary tuning and matching elements on the TMPCB ($C_1$ and $C_2$).
The lower frequency channel is connected to the detector through a high-frequency band reject ($C_7$, $L_7$).
Tuning is achieved by $L_1$ and $C_4$, which are both located on the exchangeable insert (13), in combination with
the complex admittance of the detector at the X channel frequency.
To maximise efficiency, $L_1$ should have as high a $Q$-factor as possible. We have
used commercially available flat copper coils of 5 mm diameter and 4.5 turns in the case of
\textsuperscript{15}N at 60 MHz. Tuning to \textsuperscript{13}C, either at 125 or at 150 MHz
could be achieved without $L_1$.

@@

B$_{0}$ field map of RO3035 PCB material clearly show better B$_{0}$ field homogeneity compare to FR4. To avoid the inhomogeneities in FR4 and therefore to achieve better resolution,
detectors were made from RO3035. While RO3035 provides better
resolution, it is more expensive and takes longer to obtain compared to FR4. It
therefore makes sense to first prototype new detector designs in FR4, and only
use the more expensive RO3035 once the design has been validated.

\begin{figure} \centering
\includegraphics[width=.8\linewidth,keepaspectratio=true]{./figures/pcb-fieldmap-colorbar.png}
\caption{B\textsubscript{0} field map of the sample chamber for FR4 (A) and
Rogers RO3035 (B) PCB material. Both the field maps were constructed from 2
gradient echo images acquired on a chip filled with 130 mM sodium acetate
dissolved in H\textsubscript{2}O through same parameters and pulse sequence.}
\label{fig:field map}
\end{figure}

@@

\begin{figure} \centering
\includegraphics[width=\linewidth,keepaspectratio=true]{./figures/ms5n17-tlp-sp-180612-HSQCspect.png}
\caption{ $^1$H-$^{13}$C HSQC spectrum of $^{13}$C labeled glucose (left) and
$^1$H-$^{15}$N HSQC of $^{15}$N labeled Ubiquitin (right) both acquired from
2.5$\mu$l sample volume at 14.1 T. Glucose HSQC spectrum was acquired in 12
minutes from a 100 mM sample. Ubiquitin HSQC spectrum was acquired in 400
minutes from a 1 mM sample.} \label{fig:HSQC} \end{figure}
%$
The probe can also be used to acquire micro magnetic resonance (MR) images using
a Bruker micro-gradient unit. To this effect, the aluminium probe sleeve must be
replaced by an electrically insulating material in order to avoid image
distortions due to eddy currents in the sleeve. \fig{fig:tisli} shows MR images
of mouse liver tissue slice immersed in DMEM cell growth medium. The slice was
placed in a microfluidic device with circular sample chamber centred at the
detector location, as shown in \fig{fig:tisli}(A). The image shown in
\fig{fig:tisli}(B) was acquired using a FLASH pulse sequence with $t_R=120$~ms
and $t_E=15.3$~ms. The slice thickness of  1~mm encompasses the entire thickness
of the tissue; the lateral dimensions of all MR images in \fig{fig:tisli} are
512$\times$512 pixels at a resolution of 30~$\mu$m. \fig{fig:tisli}(C) shows a
RARE image using the same parameters, with 8 echos acquired per scan, thus
reducing the acquisition time to 8 minutes. A spin echo image is shown in
\fig{fig:tisli}(D), acquired over 25~min. Finally, \fig{fig:tisli}(E) shows a
$B_0$ field map, which has been obtained by acquiring two separate FLASH images
with echo times $t_E=5.3$~ms and 15.3~ms, respectively.
